---
title: "Correlating Like a Data Master"
format: html
editor: visual
author: "Himanshu Nimbarte"
---

# **Correlating Like a Data Master**

## Setup

```{r}
if (!require(pacman))
  install.packages("pacman")

pacman::p_load(colorblindr,
       dlookr,
       formattable,
       GGally,
       ggdist,
       ggpubr,
       ggridges,
       here,
       tidyverse)

# Set global ggplot() theme
# Theme pub_clean() from the ggpubr package with base text size = 16
theme_set(theme_pubclean(base_size = 12)) 
# All axes titles to their respective far right sides
theme_update(axis.title = element_text(hjust = 1))
# Remove axes ticks
theme_update(axis.ticks = element_blank()) 
# Remove legend key
theme_update(legend.key = element_blank())
```

## Load and Examine Data

```{r}
tornados <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-05-16/tornados.csv')


tornados |>
  head() |>
  formattable()
```

## Describe and Visualize Correlation

Correlation in statistics refers to a statistical measure used to quantify the strength and direction of the linear relationship between two or more variables. It helps us understand how changes in one variable are associated with changes in another variable. Correlation is commonly used in data analysis to explore the relationships between variables and to assess the degree to which they move together or in opposite directions.

```{r}
# Table of correlations between numerical variables (we are sticking to the default Pearson's r coefficient)
correlate(tornados) |>
  formattable()
```

```{r}
# Correlation matrix of numerical variables
tornados |>
plot_correlate()
```

Correlation Matrix for attributes of the dataset

```{r}
tornados |>
  group_by(st) |>
  plot_correlate()
```

Correlation matrix based on "st"

```{r}
tornados |> 
  dplyr::select(fc,fat,loss) |>
  ggpairs(aes(color = fc, alpha = 0.5)) +
  theme(strip.background = element_blank())
```

Co-relation between FC and fat and loss

## **Describe and Visualize Relationships Based on Target Variables**

## Target Variables

"Target variable" in R typically refers to the variable that you are trying to predict or explain. It is also known as the dependent variable, response variable, or outcome variable. In regression analysis, the target variable is a continuous variable, and in classification analysis, it is a categorical variable.

### **Numerical Target Variables: Numerical Variable of Interest**

```{r}
# First, we need to remove NAs, they cause an error
tornados.noNA <- tornados |> 
  drop_na()

# The numerical predictor variable that we want
num <- target_by(tornados.noNA, fat)

# Relating the variable of interest to the numerical target variable
num_num <- relate(num,loss)

# Summary of the regression analysis - the same as the summary from lm(Formula)
summary(num_num)
```

```{r}
# Plotting the linear relationship
plot(num_num)
```

Figure shows the linear relationship between fatalities and loss

### **Numerical Target Variables: Categorical Variable of Interest**

```{r}
# The categorical predictor variable that we want
num <- target_by(tornados, fat) 

# We need to change Group to a factor
num$st <- as.factor(num$st)

# Relating the variable of interest to the numerical target variable
num_cat <- relate(num, st)

# Summary of the ANOVA analysis - the same as the summary from anova(lm(Formula))
summary(num_cat)
```

```{r}
plot(num_cat) + 
  theme(axis.text.x = element_blank())
```

Fatalities Box Plot by State

### **Categorical Target Variables: Numerical Variable of Interest**

```{r}
# The categorical predictor variable that we want
categ <- target_by(tornados.noNA, st)

# Relating the variable of interest to the numerical target variable
cat_num <- relate(categ, fat)

# Summary of descriptive statistics
summary(cat_num)
```

```{r}
plot(cat_num) 
```

States density plot by fatalities

### **Categorical Target Variables: Categorical Variable of Interest**

```{r}
# Create new categorical column
cat_dataset <- tornados |>
  select(fat, st) |>
  drop_na() |>
  mutate(HighLWP = ifelse(
    fat > (mean(fat + sd(fat))), 
                          "Yes", 
                          "No"))

# New dataset 
cat_dataset |>
  head() |>
  formattable()
```

Formula = `HighLWP (categorical) ~ Group (categorical response)`

```{r}
# The categorical predictor variable that we want
categ <- target_by(cat_dataset, HighLWP)

# Relating the variable of interest to the categorical target variable
cat_cat <- relate(categ, st)

# Summary of the Chi-square test for Independence
summary(cat_cat)
```

```{r}
plot(cat_cat)
```

Summary of chi-square test
